# Notation and naming {#style-notnam}

## File names

File names should be meaningful and end in `.R`.

    # Good
    fit_models.R
    utility_functions.R

    # Bad
    foo.r
    stuff.r

If files need to be run in sequence, prefix them with numbers:

    0_download.R
    1_parse.R
    2_explore.R

Pay attention to capitalization, since you, or some of your collaborators, might be using an operating system with a case-insensitive file system (e.g., Microsoft Windows or OS X) which can lead to problems with (case-sensitive) revision control systems. Never use filenames that differ only in capitalization.

## Object names

> "There are only two hard things in Computer Science: cache invalidation and 
> naming things." 
>
> --- Phil Karlton

Variable and function names should be lowercase. Use an underscore (`_`) to separate words within a name. Generally, variable names should be nouns and function names should be verbs. Strive for names that are concise and meaningful (this is not easy!).

Although standard R uses dots extensively in function names (`contrib.url()`), methods (`all.equal`), or class names (`data.frame`), it's better to use underscores. For example, the basic S3 scheme to define a method for a class, using a generic function, would be to concatenate them with a dot, like this `generic.class`. This can lead to confusing methods like `as.data.frame.data.frame()` whereas something like `print.my_class()` is unambiguous.

```{r, eval = FALSE}
# Good
day_one
day_1

# Bad
first_day_of_the_month
DayOne
dayone
djm1
```

Where possible, avoid using names of existing functions and variables. This will cause confusion for the readers of your code.

```{r, eval = FALSE}
# Bad
T <- FALSE
c <- 10
mean <- function(x) sum(x)
```
