# Naming {#style-notnam}

## File names

File names should be meaningful and end in `.R`. Avoid using special characters in file names - stick with numbers, letters, `-`, and `_`.


    # Good
    fit_models.R
    utility_functions.R

    # Bad
    fit models.R
    foo.r
    stuff.r

If files should be run in a particular order, prefix them with numbers:

    0_download.R
    1_parse.R
    2_explore.R

Pay attention to capitalization, since you, or some of your collaborators, might be using an operating system with a case-insensitive file system (e.g., Microsoft Windows or OS X) which can lead to problems with (case-sensitive) revision control systems. Prefer file names that are all lower case, and never have names that differ only in their capitalization.

## Object names

> "There are only two hard things in Computer Science: cache invalidation and 
> naming things." 
>
> --- Phil Karlton

Variable and function names should use only lowercase letters, numbers, and `_`. 

Use underscores (`_`) to separate words within a name. Base R uses dots in function names (`contrib.url()`) and class names (`data.frame`), but it's better to reserve dots exclusively for S3. 

Generally, variable names should be nouns and function names should be verbs. Strive for names that are concise and meaningful (this is not easy!).

```{r, eval = FALSE}
# Good
day_one
day_1

# Bad
first_day_of_the_month
DayOne
dayone
djm1
```

Where possible, avoid re-using names of common functions and variables. This will cause confusion for the readers of your code.

```{r, eval = FALSE}
# Bad
T <- FALSE
c <- 10
mean <- function(x) sum(x)
```
