# Syntax {#style-syn}

## Spacing

Place spaces around all infix operators (`=`, `+`, `-`, `<-`, etc.). The same rule applies when using `=` in function calls. Always put a space after a comma, and never before (just like in regular English).

```{r, eval = FALSE}
# Good
average <- mean(feet / 12 + inches, na.rm = TRUE)

# Bad
average<-mean(feet/12+inches,na.rm=TRUE)
```

There's a small exception to this rule: `:`, `::` and `:::` don't need spaces around them.

```{r, eval = FALSE}
# Good
x <- 1:10
base::get

# Bad
x <- 1 : 10
base :: get
```

Place a space before left parentheses, except in a function call.

```{r, eval = FALSE}
# Good
if (debug) show(x)
plot(x, y)

# Bad
if(debug)show(x)
plot (x, y)
```

Extra spacing (i.e., more than one space in a row) is ok if it improves alignment of equal signs or assignments (`<-`).

```{r, eval = FALSE}
list(
  total = a + b + c, 
  mean  = (a + b + c) / n
)
```

Do not place spaces around code in parentheses or square brackets (unless there's a comma, in which case see above).

```{r, eval = FALSE}
# Good
if (debug) do(x)
diamonds[5, ]

# Bad
if ( debug ) do(x)  # No spaces around debug
x[1,]   # Needs a space after the comma
x[1 ,]  # Space goes after comma not before
```

## Curly braces

An opening curly brace should never go on its own line and should always be followed by a new line. A closing curly brace should always go on its own line, unless it's followed by `else` (or a closing parenthesis).

Always indent the code inside curly braces. When indenting your code, use two spaces. Don't ise tabs.

```{r, eval = FALSE}
# Good
if (y < 0 && debug) {
  message("y is negative")
}

if (y == 0) {
  if (x > 0) {
    log(x)
  } else {
    message("x is negative or zero")
  }
} else {
  y ^ x
}

# Bad
if (y < 0 && debug)
message("Y is negative")

if (y == 0)
{
    if (x > 0) {
⇥       log(x)
    } else {
⇥       message("x is negative or zero")
    }
} 
else { y ^ x }
```

## Line length

Strive to limit your code to 80 characters per line. This fits comfortably on a printed page with a reasonably sized font. If you find yourself running out of room, this is a good indication that you should encapsulate some of the work in a separate function.

## Indentation

If a function call is too long, put the function name, each argument, and the closing parenthesis on a separate line. This makes the code easier to read and to change later. You may also place several arguments on the same line if they are closely related to each other, e.g., strings in calls to `paste()` or `stop()`:

```{r, eval = FALSE}
# Good
do_something_very_complicated(
  "that",
  requires = many,
  arguments = "some of which may be long"
)

# Good
paste0(
  "Requirement: ", requires, "\n",
  "Result: ", result, "\n"
)

# Bad
do_something_very_complicated("that", requires, many, arguments,
                              "some of which may be long"
                              )

# Bad
paste0(
  "Requirement: ", requires,
  "\n", "Result: ",
  result, "\n")
```

## Assignment

Use `<-`, not `=`, for assignment.

```{r}
# Good
x <- 5

# Bad
x = 5
```

## Quotes

Use `"`, not `'`, for quoting text. The only exception is when the text already contains double quotes and no single quotes.

```{r}
# Good
"Text"
'Text with "quotes"'
'<a href="http://style.tidyverse.org">A link</a>'

# Bad
'Text'
'Text with "double" and \'single\' quotes'
```

