# Syntax

## Object names

> "There are only two hard things in Computer Science: cache invalidation and 
> naming things." 
>
> --- Phil Karlton

Variable and function names should use only lowercase letters, numbers, and `_`. 

Use underscores (`_`) to separate words within a name. Base R uses dots in 
function names (`contrib.url()`) and class names (`data.frame`), but it's 
better to reserve dots exclusively for the S3 object system. In S3, methods are 
given the name `function.class`; if you also use `.` in function and class 
names you end up with confusing methods like `as.data.frame.data.frame()`.

Generally, variable names should be nouns and function names should be verbs. 
Strive for names that are concise and meaningful (this is not easy!).

```{r, eval = FALSE}
# Good
day_one
day_1

# Bad
first_day_of_the_month
DayOne
dayone
djm1
```

Where possible, avoid re-using names of common functions and variables. This 
will cause confusion for the readers of your code.

```{r, eval = FALSE}
# Bad
T <- FALSE
c <- 10
mean <- function(x) sum(x)
```

## Spacing

Put a space before and after `=` when naming arguments in function calls.
Most infix operators (`==`, `+`, `-`, `<-`, etc.) are also surrounded by
spaces, except those with relatively high precedence: `^`, `:`, `::`, and `:::`.

Always put a space after a comma,
and never before (just like in regular English).  Prefer parentheses over
irregular spacing to highlight operator precedence.

```{r, eval = FALSE}
# Good
average <- mean((feet / 12) + inches, na.rm = TRUE)
sqrt(x^2 + y^2)
x <- 1:10
base::get

# Bad
average<-mean(feet/12 + inches,na.rm=TRUE)
sqrt(x ^ 2 + y ^ 2)
x <- 1 : 10
base :: get
```

Place a space before `(`, except when it's part of a function call.

```{r, eval = FALSE}
# Good
if (debug) show(x)
plot(x, y)

# Bad
if(debug)show(x)
plot (x, y)
```

Extra spacing (i.e., more than one space in a row) is ok if it improves 
alignment of equal signs or assignments (`<-`).

```{r, eval = FALSE}
list(
  total = a + b + c,
  mean  = (a + b + c) / n
)
```

Do not place spaces around code in parentheses or square brackets (unless 
there's a comma, in which case see above).

```{r, eval = FALSE}
# Good
if (debug) do(x)
diamonds[5, ]

# Bad
if ( debug ) do(x)  # No spaces around debug
x[1,]   # Needs a space after the comma
x[1 ,]  # Space goes after comma not before
```

## Argument names

A function's arguments typically fall into two broad categories: one supplies 
the __data__ to compute on; the other controls __details__ of computation. When 
you call a function, you typically omit the names of data arguments, because 
they are used so commonly. If you override the default value of an argument, 
use the full name:

```{r, eval = FALSE}
# Good
mean(1:10, na.rm = TRUE)

# Bad
mean(x = 1:10, , FALSE)
mean(, TRUE, x = c(1:10, NA))
```

Avoid partial matching.

## Indenting

Curly braces, `{}`, define the the most important hierarchy of R code. To make 
this hierarchy easy to see, always indent the code inside `{}` by two spaces.

A symmetrical arrangement helps finding related braces: The opening brace is the
last, the closing brace is the first non-whitespace character in a line.  Code
that is related to a brace (e.g., an `if` clause, a function declaration, a
trailing comma, ...) must be on the same line.

```{r, eval = FALSE}
# Good
if (y < 0 && debug) {
  message("y is negative")
}

if (y == 0) {
  if (x > 0) {
    log(x)
  } else {
    message("x is negative or zero")
  }
} else {
  y ^ x
}

test_that("call1 returns an ordered factor", {
  expect_s3_class(call1(x, y), c("factor", "ordered"))
})

tryCatch(
  {
    x <- scan()
    cat("Total: ", sum(x), "\n", sep = "")
  },
  interrupt = function(e) {
    message("Aborted by user")
  }
)

# Bad
if (y < 0 && debug)
message("Y is negative")

if (y == 0)
{
    if (x > 0) {
      log(x)
    } else {
  message("x is negative or zero")
    }
} else { y ^ x }
```

It's ok to drop the curly braces if you have a very simple and short `if` 
statement that fits on one line. If you have any doubt, it's better to use the 
full form.

```{r}
y <- 10
x <- if (y < 20) "Too low" else "Too high"
```

## Long lines

Strive to limit your code to 80 characters per line. This fits comfortably on a 
printed page with a reasonably sized font. If you find yourself running out of 
room, this is a good indication that you should encapsulate some of the work in 
a separate function.
If a function call is too long to fit on a single line, use one line each for 
the function name, each argument, and the closing `)`. 
This makes the code easier to read and to change later. 
```{r, eval = FALSE}
# Good
do_something_very_complicated(
  something = "that",
  requires = many,
  arguments = "some of which may be long"
)

# Bad
do_something_very_complicated("that", requires, many, arguments,
                              "some of which may be long"
                              )
```

As described under [Argument names], you can omit the argument names
for very common arguments (i.e. for arguments that are used in almost every 
invocation of the function). Short unnamed arguments can also go on the same 
line as the function name, even if the whole function call spans multiple lines.
```{r, eval = FALSE}
map(x, f,
  extra_argument_a = 10,
  extra_argument_b = c(1, 43, 390, 210209)
)
```

You may also place several arguments on the same line if they are closely 
related to each other, e.g., strings in calls to `paste()` or `stop()`. When 
building strings, where possible match one line of code to one line of output. 

```{r, eval = FALSE}
# Good
paste0(
  "Requirement: ", requires, "\n",
  "Result: ", result, "\n"
)

# Bad
paste0(
  "Requirement: ", requires,
  "\n", "Result: ",
  result, "\n")
```

## Assignment

Use `<-`, not `=`, for assignment.

```{r}
# Good
x <- 5

# Bad
x = 5
```

## Semicolons

Don't put `;` at the end of a line, and don't use `;` to put multiple commands 
on one line.

## Quotes

Use `"`, not `'`, for quoting text. The only exception is when the text already 
contains double quotes and no single quotes.

```{r}
# Good
"Text"
'Text with "quotes"'
'<a href="http://style.tidyverse.org">A link</a>'

# Bad
'Text'
'Text with "double" and \'single\' quotes'
```

