# Syntax

## Object names

> "There are only two hard things in Computer Science: cache invalidation and 
> naming things." 
>
> --- Phil Karlton

Variable and function names should use only lowercase letters, numbers, and `_`. 

Use underscores (`_`) to separate words within a name. Base R uses dots in 
function names (`contrib.url()`) and class names (`data.frame`), but it's 
better to reserve dots exclusively for the S3 object system. In S3, methods are 
given the name `function.class`; if you also use `.` in function and class 
names you end up with confusing methods like `as.data.frame.data.frame()`.

Generally, variable names should be nouns and function names should be verbs. 
Strive for names that are concise and meaningful (this is not easy!).

```{r, eval = FALSE}
# Good
day_one
day_1

# Bad
first_day_of_the_month
DayOne
dayone
djm1
```

Where possible, avoid re-using names of common functions and variables. This 
will cause confusion for the readers of your code.

```{r, eval = FALSE}
# Bad
T <- FALSE
c <- 10
mean <- function(x) sum(x)
```

## Spacing

Put a space before and after all infix operators (=, +, -, <-, etc.), and when 
naming arguments in function calls. Always put a space after a comma, and never 
before (just like in regular English).

```{r, eval = FALSE}
# Good
average <- mean(feet / 12 + inches, na.rm = TRUE)

# Bad
average<-mean(feet/12+inches,na.rm=TRUE)
```

There's a small exception to this rule: don't use spaces around `:`, `::`, and 
`:::`.

```{r, eval = FALSE}
# Good
x <- 1:10
base::get

# Bad
x <- 1 : 10
base :: get
```

Place a space before `(`, except when it's part of a function call.

```{r, eval = FALSE}
# Good
if (debug) show(x)
plot(x, y)

# Bad
if(debug)show(x)
plot (x, y)
```

Extra spacing (i.e., more than one space in a row) is ok if it improves 
alignment of equal signs or assignments (`<-`).

```{r, eval = FALSE}
list(
  total = a + b + c,
  mean  = (a + b + c) / n
)
```

Do not place spaces around code in parentheses or square brackets (unless 
there's a comma, in which case see above).

```{r, eval = FALSE}
# Good
if (debug) do(x)
diamonds[5, ]

# Bad
if ( debug ) do(x)  # No spaces around debug
x[1,]   # Needs a space after the comma
x[1 ,]  # Space goes after comma not before
```

## Argument names

A function's arguments typically fall into two broad categories: one supplies 
the __data__ to compute on; the other controls __details__ of computation. When 
you call a function, you typically omit the names of data arguments, because 
they are used so commonly. If you override the default value of an argument, 
use the full name:

```{r, eval = FALSE}
# Good
mean(1:10, na.rm = TRUE)

# Bad
mean(x = 1:10, , FALSE)
mean(, TRUE, x = c(1:10, NA))
```

Avoid partial matching.

## Indenting

Curly braces, `{}`, define the the most important hierarchy of R code. To make 
this hierarchy easy to see, always indent the code inside `{}` by two spaces. 
A `{` should never go on its own line and should always be followed by a new 
line, unless it is part of a free-standing block. A `}` should always go on its
own line, unless it's followed by `else`, `,` or `)`. 

```{r, eval = FALSE}
# Good
if (y < 0 && debug) {
  message("y is negative")
}

if (y == 0) {
  if (x > 0) {
    log(x)
  } else {
    message("x is negative or zero")
  }
} else {
  y ^ x
}

test_that("call1 returns an ordered factor", {
  expect_s3_class(call1(x, y), c("factor", "ordered"))
})

tryCatch(
  {
    x <- scan()
    cat("Total: ", sum(x), "\n", sep = "")
  },
  interrupt = function(e) {
    message("Aborted by user")
  }
)

# Bad
if (y < 0 && debug)
message("Y is negative")

if (y == 0)
{
    if (x > 0) {
      log(x)
    } else {
  message("x is negative or zero")
    }
} else { y ^ x }
```

It's ok to drop the curly braces if you have a very simple and short `if` 
statement that fits on one line. If you have any doubt, it's better to use the 
full form.

```{r}
y <- 10
x <- if (y < 20) "Too low" else "Too high"
```

## Long lines

Strive to limit your code to 80 characters per line. This fits comfortably on a 
printed page with a reasonably sized font. If you find yourself running out of 
room, this is a good indication that you should encapsulate some of the work in 
a separate function.

If a function call is too long to fit on a single line, use one line each for 
the function name, each argument, and the closing `)`. This makes the code 
easier to read and to change later. 

```{r, eval = FALSE}
# Good
do_something_very_complicated(
  "that",
  requires = many,
  arguments = "some of which may be long"
)

# Bad
do_something_very_complicated("that", requires, many, arguments,
                              "some of which may be long"
                              )
```

You may also place several arguments on the same line if they are closely 
related to each other, e.g., strings in calls to `paste()` or `stop()`. When 
building strings, where possible match one line of code to one line of output. 

```{r, eval = FALSE}
# Good
paste0(
  "Requirement: ", requires, "\n",
  "Result: ", result, "\n"
)

# Bad
paste0(
  "Requirement: ", requires,
  "\n", "Result: ",
  result, "\n")
```

## Assignment

Use `<-`, not `=`, for assignment.

```{r}
# Good
x <- 5

# Bad
x = 5
```

## Semicolons

Don't put `;` at the end of a line, and don't use `;` to put multiple commands 
on one line.

## Quotes

Use `"`, not `'`, for quoting text. The only exception is when the text already 
contains double quotes and no single quotes.

```{r}
# Good
"Text"
'Text with "quotes"'
'<a href="http://style.tidyverse.org">A link</a>'

# Bad
'Text'
'Text with "double" and \'single\' quotes'
```

