# Functions

## Naming

Use verbs for function names, where possible.

```{r eval = FALSE}
# Good
add_row()
permute()

# Bad
row_adder()
permutation()
```

## Size

Strive to keep blocks within a function on one screen, <30 lines.

## Long lines

If a function definition runs over multiple lines, indent the second line to where the definition starts.

```{r, eval = FALSE}
# Good
long_function_name <- function(a = "a long argument", 
                               b = "another argument",
                               c = "another long argument") {
  # As usual code is indented by two spaces.
}

# Bad
long_function_name <- function(a = "a long argument",
  b = "another argument",
  c = "another long argument") {
  # Here it's hard to spot where the definition end and the
  # code begins
}
```

## `return()`

Only use `return()` for early returns. Otherwise rely on R to return the result of the last evaluted expression.

```{r eval = FALSE}
# Good
find_abs <- function(x, y){
  if (x > 0) return(x)
  x * -1
}

# Bad
add_two <- function(x, y) {
  return(x + y)
}
```

## Comments

Use comments to explain the "why" not the "what" or "how". Each line of a comment should begin with the comment symbol and a single space: `# `.

