# Functions

## Naming

Use verbs for function names, where possible.

```{r eval = FALSE}
# Good
add_row()
permute()

# Bad
row_adder()
permutation()
```

## Long lines

If a function definition runs over multiple lines, indent the second line to where the definition starts.

```{r, eval = FALSE}
# Good
long_function_name <- function(a = "a long argument",
                               b = "another argument",
                               c = "another long argument") {
  # As usual code is indented by two spaces.
}

# Bad
long_function_name <- function(a = "a long argument",
  b = "another argument",
  c = "another long argument") {
  # Here it's hard to spot where the definition end and the
  # code begins
}
```

## `return()`

Only use `return()` for early returns. Otherwise rely on R to return the result of the last evaluted expression.

```{r eval = FALSE}
# Good
find_abs <- function(x, y){
  if (x > 0) return(x)
  x * -1
}

# Bad
add_two <- function(x, y) {
  return(x + y)
}
```

If your function is called primarily for its side-effects, it should return the first argument invisibly. This makes it possible to use the function as part of the pipe. `print` methods should usually do this, like this example from httr:

```{r eval = FALSE}
print.url <- function(x, ...) {
  cat("Url: ", build_url(x), "\n", sep = "")
  invisible(x)
}
```


## Comments

Use comments to explain the "why" not the "what" or "how". Each line of a comment should begin with the comment symbol and a single space: `# `.

## Design

There are two main design principles to bear in mind:

*   A function should do one thing well. 
 
    A function should be called because it has side-effects or because it
    returns a value; not both. Strive to keep blocks within a function on 
    one screen, <30 lines.

*   A function should be easily understandable in isolation.

    Avoid global options. If your function has a transient side-effect
    (i.e. you need to create a temporary file or set an option), clean up
    after yourself with `on.exit()`.
