# Pipes {#style-pipes}

Use the `%>%` operator from the tidyverse when you find yourself composing more than two functions together into a nested call, or when you find yourself creating unnecessaty interim objects to avoid nesting functions together.

```{r eval = FALSE}
# Good
foo_foo %>%
  hop(through = forest) %>%
  scoop(up = field_mouse) %>%
  bop(on = head)

# Bad
bop(
  scoop(
    hop(foo_foo, through = forest),
    up = field_mice
  ), 
  on = head
)

foo_foo <- hop(foo_foo, through = forest)
foo_foo <- scoop(foo_foo, up = field_mice)
foo_foo <- bop(foo_foo, on = head)
```

When you use the `%>%` operator, put each verb on its own line.  This makes it simpler to rearrange them later, and makes it harder to overlook a step. It is ok to keep a one-step pipe in one line.

```{r, eval = FALSE}
# Good
iris %>%
  group_by(Species) %>%
  summarize_all(mean) %>%
  ungroup %>%
  gather(measure, value, -Species) %>%
  arrange(value)

iris %>% arrange(Petal.Width)

# Bad
iris %>% group_by(Species) %>% summarize_all(mean) %>%
  ungroup %>% gather(measure, value, -Species) %>%
  arrange(value)
```

## When not to use the pipe

Do not use a pipe if:

* You need to manipulate more than one object at a time. Reserve pipes for a sequence of steps applied to one primary object.

* You have more than (say) ten steps in your pipe. In that case, divide your pipe into shorter pipes that create intermediate objects with meaningful names.
