# Pipes

## Introduction

Use `%>%` when you find yourself composing more than three or more functions together into a nested call, or creating intermediate objects that you don't care about.

```{r eval = FALSE}
# Good
foo_foo %>%
  hop(through = forest) %>%
  scoop(up = field_mouse) %>%
  bop(on = head)

# Bad
foo_foo <- hop(foo_foo, through = forest)
foo_foo <- scoop(foo_foo, up = field_mice)
foo_foo <- bop(foo_foo, on = head)

# Bad
bop(
  scoop(
    hop(foo_foo, through = forest),
    up = field_mice
  ),
  on = head
)
```

Avoid using the pipe when:

* You need to manipulate more than one object at a time. Reserve pipes for a 
  sequence of steps applied to one primary object.
 
* You there are meaningful intermediate objects that could be given
  informative names.

## Spacing and indenting

`%>%` should always have a space before it and a new line after it. After the first step, each line should be indented by two spaces.

```{r, eval = FALSE}
# Good
iris %>%
  group_by(Species) %>%
  summarize_if(is.numeric(), mean) %>%
  ungroup %>%
  gather(measure, value, -Species) %>%
  arrange(value)

# Bad
iris %>% group_by(Species) %>% summarize_all(mean) %>%
ungroup %>% gather(measure, value, -Species) %>%
arrange(value)
```

It is ok to keep a one-step pipe on one line, although depending on the context, you may want to rewrite to remove the pipe.

```{r, eval = FALSE}
# good
iris %>% arrange(Petal.Width)
arrange(iris, Petal.Width)
```

## Long lines

If the arguments to a function don't all fit on one line, put each argument on its own line and indent:

```{r, eval = FALSE}
iris %>%
  group_by(Species) %>%
  summarise(
    Sepal.Length = mean(Sepal.Length),
    Sepal.Width = mean(Sepal.Width),
    Sepcies = n_distinct(Species)
  )
```

## Assignment

You can use to use `->` to create an object at the end of the pipe.

Personally, I think you should avoid this technique. While starting with assignment is a little more work when writing, but makes it easier to read the code because the name acts as a heading, reminding you of the purpose of the code.

## Comments

Use comments to record important data analysis findings.  If you discover that you have more comments than code, considering switching to RMarkdown.

