# Code Documentation

## Introduction

Documentation of code is essential, even if the only person using your code 
is future-you. Use 
[roxygen2](https://github.com/klutometis/roxygen) with enabled [markdown](https://github.com/klutometis/roxygen/blob/master/vignettes/markdown.Rmd)
support to keep the documentation close to your code.

```{r, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Title and Description
For the title, describe concisely what the function does at the very 
top of your function documentation. Do not use a dot at the end of the title 
and use title style capitalization. For both the title and the description, 
there is no need to use the explicit `@title` or `@description` tag. To 
visually separate the title and the description, leave one line blank after 
each of them.
```{r}
#' Combine Values into a Vector or List
#' 
#' This is a generic function which combines its arguments.
#'
```

## Indention
Always indent with one space after `#'`. If any description corresponding to a 
`roxygen` tag spans over multiple lines, add another two spaces of extra 
indention.
```{r}
#' @param key The bare (unquoted) name of the column whose values will be used 
#'   as column headings. 
```

Alternatively, tags that span over multiple lines (like `@examples`) can have 
the corresponding tag on its own line and subsequent lines do not need extra
indention.
```{r}
#' @examples
#' 1 + 1
#' sin(pi)
```
The section 'Description' does not need extra indention either.

## Capitalization
The first letter of every description corresponding to a `roxygen` tag is 
upper case.
```{r}
#' @param key The bare (unquoted) name of the column whose values will be used 
#'   as column headings. 
```

An exception to this rule are words in `code font` (see below), in which case
also lower case letters can occur.
```{r}
#' @seealso [fct_lump()] 
```

## Periods
For all bullets, enumerations argument descriptions and the like, put a period 
at the end of each text element, even if it is only a few words. Use a double 
point before you start to enumerate.
```{r}
#' @details In the following, we present the bullets of the list:
#' * item one.
```

## Reference
Cross-referencing is encouraged, both within R's help file system as well as to 
external resources. Use `()` when referencing function calls after the function
name.
```{r}
#' @seealso [fct_lump()] 
```

When referring to other sections in the documentation, use single quotes and 
upper title style capitalization.
```{r}
#' @param p1 Another parameter. See 'Details'.
```

## Line Break
To force a line break in the `.Rd` file, simply put an empty line after the 
line you want to insert the line break and `roxygen2` will do the job.
```{r}
#' Some stuff.
#' 
#' Some remotely related stuff on a new line.
```

## `Code Font`
As a general rule, use code font when you are referring to something that has a 
specific and well-defined meaning within the R language or your code 
specifically. If such a specific meaning is attributed to a term, but you mean 
just the general concept, do not use code font.
```{r}
#' A matrix can store data. Matematicians used matrices before R existed.
```

The general rule introduced above implies to use `code font` for:

* Function names.
* Function arguments. 
* Logical values.
* Literal R code.
* `NA`, `NaN`, `...`, `NULL` etc.

To refer to the R language, use \\R.