# (PART) Packages {-}

# Documentation

## Introduction

Documentation of code is essential, even if the only person using your code 
is future-you. Use [roxygen2](https://github.com/klutometis/roxygen) with 
[markdown](https://github.com/klutometis/roxygen/blob/master/vignettes/markdown.Rmd)
support enabled to keep your documentation close to the code.

```{r, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

## Title and description

Use the first line of your function documentation to provide a concise title that describes the function, dataset, or class. Titles should use sentence case
but not end with a full stop (`.`).

```{r}
#' Combine values into a vector or list
#' 
#' This is a generic function which combines its arguments.
#'
```

There is no need to use the explicit `@title` or `@description` tags, except 
in the case of the description if it is multiple paragraphs or includes 
more complex formatting like a bulleted list.

```{r}
#' Apply a function to each element of a vector
#'
#' @description
#' The map function transform the input, returning a vector the same length
#' as the input. 
#' 
#' * `map()` returns a list or a data frame
#' * `map_lgl()`, `map_int()`, `map_dbl()` and `map_chr()` return 
#'     vectors of the corresponding type (or die trying); 
#' * `map_dfr()` and `map_dfc()` return data frames created by row-binding 
#'    and column-binding respectively. They require dplyr to be installed.
```

## Indents and line breaks

Always indent with one space after `#'`. If any description corresponding to a 
`roxygen` tag spans over multiple lines, add another two spaces of extra 
indention.

```{r}
#' @param key The bare (unquoted) name of the column whose values will be used 
#'   as column headings. 
```

Alternatively, tags that span over multiple lines (like `@description`, `@examples` and `@section`) can have the corresponding tag on its own line and then subsequent lines don't need to be indented.

```{r}
#' @examples
#' 1 + 1
#' sin(pi)
```

Use line breaks before/after sections where needed:

```{r}
#' @section Tidy data:
#' When applied to a data frame, row names are silently dropped. To preserve,
#' convert to an explicit variable with [tibble::rownames_to_column()].
#'
#' @section Scoped filtering:
#' The three [scoped] variants ([filter_all()], [filter_if()] and
#' [filter_at()]) make it easy to apply a filtering condition to a
#' selection of variables.
```

## Documenting parameters

For most tags, like `@param`, `@seealso` and `@return`, the text should be a 
sentence, starting with a capital letter and ending with a full stop.

```{r}
#' @param key The bare (unquoted) name of the column whose values will be used 
#'   as column headings. 
```

If some functions share parameters, you can use `@inheritParams` to avoid 
duplication of content in multiple places.

```{r}
#' @inheritParams argument function_to_inherit_from
```

## Capitalization and full stops

For all bullets, enumerations, argument descriptions and the like, use sentence
case and put a period at the end of each text element, even if it is only a few 
words. However, avoid capitalization of function names or packages since R is 
case sensitive. Use a colon before enumerations or bulleted lists.

```{r}
#' @details 
#' In the following, we present the bullets of the list:
#' * Four cats are few animals.
#' * forcats is a package.
```

## Cross-linking

Cross-referencing is encouraged, both within R's help file system as well as to 
external resources. Include parentheses after function names when referencing 
function calls.

```{r}
#' @seealso [fct_lump()] 
```

If you have a family of related functions, you can use the `@family` tag to 
automatically add appropriate lists and interlinks to the `@seealso` section.
Family names are plural. In dplyr, the verbs `arrange()`, `filter()`, 
`mutate()`, `slice()`, `summarize()` form the family of single table verbs.

```{r}
#' @family single table verbs
```

## R code

Text that contains valid R code should be marked as such using backticks. This includes:

* Function names, which should be followed by `()`, e.g. `tibble()`.
* Function arguments, e.g. `na.rm`. 
* Values, e.g. `TRUE`, `FALSE`, `NA`, `NaN`, `...`, `NULL` 
* Literal R code, e.g. `mean(x, na.rm = TRUE)`

Do not use code font for package names. If package name is ambiguous in the 
context, disambiguate with words, e.g. "the foo package".

## Internal functions

Internal functions should be documented with `#'` comments as per usual. 
Use the `@noRd` tag to prevent `.Rd` files from being generated.

```{r}
#' Drop last
#'
#' Drops the last element from a vector.
#'
#' @param x A vector object to be trimmed.
#'
#' @noRd
```
