# Object Oriented Programming

## Introduction

R has three object oriented programming (OOP) systems: S3, S4 and R6. Broadly 
speaking, the previously outlined aspects of code style still apply within each 
of these systems, with a few exceptions (e.g. in contrast to the general naming 
rules outlined in Section \@ref(object-names), S4 and R6 class names should by 
convention be UpperCamelCase).

However, as highlighted in [Advanced R](http://adv-r.hadley.nz/oo-tradeoffs.html), 
S4 should generally be avoided (since in the majority of cases the added 
complexity does not yield any significant benefits over S3), and R6 should only 
be used if there are compelling reasons to do so.

This chapter will thus focus only on S3. S3 is the most commonly used within the 
tidyverse (and CRAN packages in general) and should be the default choice for 
OOP. S3 has few built-in constraints and is very flexible. This makes it 
especially important that developers adhere to best practice and convention in 
order to use S3 sensibly.

## Class names

Class names in S3 should contain only lower case letters and underscores (i.e. 
snake_case). Dots should be strictly avoided due to the confusion this creates 
with the identification of S3 methods.

```{r, eval = FALSE}
# Good
foo_bar

# Bad
foo.bar
fooBar
FooBar
```

## Constructors

When creating an S3 class, you should always provide a constructor that 
efficiently creates new objects. The constructor should:

- be named `new_classname()`
- have an argument for the base object, and one for each attribute
- check types of the base object and each attribute

```{r, eval = FALSE}
# Good
new_foo_bar <- function(x, size = "small") {
  stopifnot(is.double(x))
  stopifnot(is.character(size), length(size) == 1)
  
  structure(x, class = "foo_bar", size = size)
}

# Bad
foo_bar_maker <- function(x) {
  structure(x$x, class = "foo_bar", size = x$size)
}
```

Constructors usually shouldn't do any additional checking, in order to keep them 
lightweight and performant.

## Validators

A validator performs more expensive checks on a created object to see that it 
has the expected properties. The validator should be named `validate_classname()` 
and should return the object if it is valid (this allows the validator to be 
re-used in the helper function and also supports piping).

```{r, eval = FALSE}
# Good
validate_foo_bar <- function(x) {
  
  values <- unclass(x)
  size <- attr(x, "size")
  
  if (!all(!is.na(values) & values > 0)) {
    stop(
      "All `x` values must be non-missing and greater than zero",
      call. = FALSE
    )
  }
  
  if (!(size %in% c("small", "medium", "large"))) {
    stop(
      "The `size` attribute needs to be either 'small', 'medium', or 'large'",
      call. = FALSE
    )
  }
  
  x
}
```

## Helpers

It's a good idea, especially for complex classes, to also provide a helper, that 
provides a convenient and neatly parameterised way for others to construct and 
validate objects of this class. This function should be called `classname()`.

```{r, eval = FALSE}
# Good
foo_bar <- function(x, size = "small") {
  validate_foo_bar(new_foo_bar(x, size = size))
}
```

## Base types

When creating classes, you should generally use a scalar (list) style, since 
this is easiest to work with. 

Otherwise it’s usually clear if you need to use a data-frame style. Only use the 
vector style if your class closely resembles an existing vector type.

## Generics and methods

Generics should generally dispatch based on the first argument (unless there is 
an existing constraint or good reason not to).

When creating a new generic, you should implement `generic.default()`.

Always avoid calling S3 methods directly, except in special circumstances when 
you can guarantee the type of the argument(s) and need to bypass method dispatch 
in order to improve performance.

## Arguments

Methods should always have the same arguments as their generics, because this 
will avoid confusing behaviour.

The one exception to this is when a generic has the `...` argument - as this 
allows methods to have additional arguments.

## Coercion

If the S3 object can be naturally created from existing objects (of certain 
types), you should provide an (S3 generic) coercion function called 
`as_classname()` and implement coercion methods for the relevant base types 
which re-use the constructor/validator/helper. It's also good practice to 
provide a default method that gives a suitable error message when coercion isn't 
possible.

Note that though class is stored as an attribute, one should never change this 
attribute directly.

```{r, eval = FALSE}
# Bad
attr(x, "class") <- "foo"
```

If the class must be set directly, use the `class()` function, since this 
communicates intent more clearly.

```{r, eval = FALSE}
# Good
class(x) <- "foo"
```
