# Error messages

(This guide is almost entirely aspirational; most of the bad examples come from existing tidyverse code.)

## Problem statement

Every error message should start with general statement of the problem. It should be concise, but informative. (This is hard!)

*   If the cause of the problem is clear use "must":

    ```{r, eval = FALSE}
    dplyr::nth(1:10, "x")
    #> Error: `n` must be a numeric vector, not a character vector
    
    dplyr::nth(1:10, 1:2)
    #> Error: `n` must have length 1, not length 2
    ```
    
    Clear cut causes typically involve incorrect types or lengths.

*   If you can't state what was expected, use "can't":

    ```{r, eval = FALSE}
    mtcars %>% pull(b)
    #> Error: Can't find column `b` in `.data`
    
    as_vector(environment())
    #> Error: Can't coerece `.x` to a vector
    
    purrr::modify_depth(list(list(x = 1)), 3, ~ . + 1)
    #> Error: Can't find specified `.depth` in `.x`
    ```

Use `stop(call. = TRUE)`, `rlang::abort()`, `Rf_errorcall(R_NilValue, ...)` to avoid cluttering the error message with the name of the function that generated it. That information is often not informative, and can easily be accessed via `traceback()` or IDE equivalent.

## Error location

Do your best to reveal the location, name, and/or content of the troublesome component. The goal is to make it easy as possible for the user to find and fix the problem.

```{r, eval = FALSE}
# GOOD
map_int(1:5, ~ "x")
#> Error: Each result must be a single integer; result 1 is a character vector

# BAD
map_int(1:5, ~ "x")
#> Error: Each result must be a single integer
```

(It is often not easy to identify the exact problem; it may require passing around extra arguments so that error messages generated at a lower-level can know the original source. For frequently used functions, the effort is typically worth it.)

If the source of the error is unclear, avoid pointing the user in the wrong direction by giving an opinion about the source of the error:

```{r, eval = FALSE}
# GOOD
pull(mtcars, b)
#> Error: Can't find column `b` in `.data`

tibble(x = 1:2, y = 1:3, z = 1)
#> Error: Columns must have consistent lengths: 
#> * Column `x` has length 2
#> * Column `y` has length 3

# BAD: implies one argument at fault
pull(mtcars, b)
#> Error: Column `b` doesn't exist in `.data`

pull(mtcars, b)
#> Error: `.data` must contain column `b`

tibble(x = 1:2, y = 1:3, z = 1)
#> Error: Column `x` must be length 1 or 3, not 2 
```

If the source of the error is clear and common, you may want provide a hint as how to fix it:

```{r}
dplyr::filter(iris, Species = "setosa")
#> Error: Filter specifications must be named. 
#> Did you mean `Species == "setosa"`?
```

If there are multiple issues, or an inconsistency revealed across several arguments or items, prefer a bulleted list:

```{r, eval = FALSE}
# GOOD
purrr::reduce2(1:4, 1:2, `+`)
#> Error: `.x` and `.y` must have compatible lengths:
#> * `.x` has length 4
#> * `.y` has length 2

# BAD: harder to scan
purrr::reduce2(1:4, 1:2, `+`)
#> Error: `.x` and `.y` must have compatible lengths: `.x` has length 4 and 
#> `.y` has length 2
```

## Punctuation

*   Errors generally follow the same rule as headings; they should be written in 
    sentence case, and should not have a full stop. 
    
*   Prefer the use of the singular:

    ```{r, eval = FALSE}
    # GOOD
    map_int(1:2, ~ "a")
    #> Error: Each result must be coercible to a single integer; result 1 is a character vector
    
    # BAD
    map_int(1:2, ~ "a")
    #> Error: Results must be a coercible to single integers; result 1 is a character vector
    ```

*   If you can detect multiple problems, list up to five. This allows the user
    to fix multiple problems in a single pass without being overwhelmed by
    many errors that may have the same source.

    ```{r}
    # BETTER
    map_int(1:10, ~ "a")
    #> Error: Results must be a coercible to single integers:
    #> * Result 1 is a character vector
    #> * Result 2 is a character vector
    #> * Result 3 is a character vector
    #> * Result 4 is a character vector
    #> * Result 5 is a character vector
    #> * ... and 5 more problems
    ```

*   Pick a natural connector between problem statement and error location:
    this may be ", not", ";", or ":" depending on the context.

*   Surround the names of variables in backticks, e.g. `` `x` ``
 
*   Ideally, each component of the error message should be less than 80 
    characters wide. Do not add manual line breaks to long error messages; 
    they will not look correct if the console is narrower (or much wider) than 
    expected.
