# Functions

## Naming

Use verbs for function names, where possible.

```{r eval = FALSE}
# Good
add_row()
permute()

# Bad
row_adder()
permutation()
```

## Long lines

If a function definition runs over multiple lines, indent the second line to 
where the definition starts.

```{r, eval = FALSE}
# Good
long_function_name <- function(a = "a long argument",
                               b = "another argument",
                               c = "another long argument") {
  # As usual code is indented by two spaces.
}

# Bad
long_function_name <- function(a = "a long argument",
  b = "another argument",
  c = "another long argument") {
  # Here it's hard to spot where the definition ends and the
  # code begins
}
```

## `return()`

Only use `return()` for early returns. Otherwise, rely on R to return the result 
of the last evaluated expression.

```{r eval = FALSE}
# Good
find_abs <- function(x, y) {
  if (x > 0) return(x)
  x * -1
}
add_two <- function(x, y) {
  x + y
}

# Bad
add_two <- function(x, y) {
  return(x + y)
}
```

If your function is called primarily for its side-effects (like printing, 
plotting, or saving to disk), it should return the first argument invisibly. 
This makes it possible to use the function as part of a pipe. `print` methods 
should usually do this, like this example from [httr](http://httr.r-lib.org/):

```{r eval = FALSE}
print.url <- function(x, ...) {
  cat("Url: ", build_url(x), "\n", sep = "")
  invisible(x)
}
```

## Comments

In code, use comments to explain the "why" not the "what" or "how". Each line 
of a comment should begin with the comment symbol and a single space: `# `.

