# Functions

## Naming

As well as following the general advice for [object names], strive to use verbs for function names:

```{r eval = FALSE}
# Good
add_row()
permute()

# Bad
row_adder()
permutation()
```

## Long lines

There are two options if the function name and definition can't fit on a single line:

*   Function-indent: place each argument on its own line, and indent to match the opening `(` of `function`:

    ```{r, eval = FALSE}
    long_function_name <- function(a = "a long argument",
                                   b = "another argument",
                                   c = "another long argument") {
      # As usual code is indented by two spaces.
    }
    ```

*   Double-indent: Place each argument of its own **double** indented line.

    ```{r, eval = FALSE}
    long_function_name <- function(
        a = "a long argument",
        b = "another argument",
        c = "another long argument") {
      # As usual code is indented by two spaces.
    }
    ```

In both cases the trailing `)` and leading `{` should go on the same line as the last argument.

Prefer function-indent style to double-indent style when it fits.

These styles are designed to clearly separate the function definition from its body.

```{r, eval = FALSE}
# Bad
long_function_name <- function(a = "a long argument",
  b = "another argument",
  c = "another long argument") {
  # Here it's hard to spot where the definition ends and the
  # code begins, and to see all three function arguments
}
```

If a function argument can't fit on a single line, this is a sign you should rework the argument to keep it [short and sweet](https://design.tidyverse.org/def-short.html).

## `return()`

Only use `return()` for early returns. Otherwise, rely on R to return the result 
of the last evaluated expression.

```{r eval = FALSE}
# Good
find_abs <- function(x) {
  if (x > 0) {
    return(x)
  }
  x * -1
}
add_two <- function(x, y) {
  x + y
}

# Bad
add_two <- function(x, y) {
  return(x + y)
}
```

Return statements should always be on their own line because they have important effects on the control flow. See also [inline statements](#inline-statements).

```{r, eval = FALSE}
# Good
find_abs <- function(x) {
  if (x > 0) {
    return(x)
  }
  x * -1
}

# Bad
find_abs <- function(x) {
  if (x > 0) return(x)
  x * -1
}
```

If your function is called primarily for its side-effects (like printing, 
plotting, or saving to disk), it should return the first argument invisibly. 
This makes it possible to use the function as part of a pipe. `print` methods 
should usually do this, like this example from [httr](http://httr.r-lib.org/):

```{r eval = FALSE}
print.url <- function(x, ...) {
  cat("Url: ", build_url(x), "\n", sep = "")
  invisible(x)
}
```

## Comments

In code, use comments to explain the "why" not the "what" or "how". Each line 
of a comment should begin with the comment symbol and a single space: `# `.

```{r, eval = FALSE}
# Good

# Objects like data frames are treated as leaves
x <- map_if(x, is_bare_list, recurse)


# Bad

# Recurse only with bare lists
x <- map_if(x, is_bare_list, recurse)
```

Comments should be in sentence case, and only end with a full stop if they
contain at least two sentences:

```{r, eval = FALSE}
# Good

# Objects like data frames are treated as leaves
x <- map_if(x, is_bare_list, recurse)

# Do not use `is.list()`. Objects like data frames must be treated
# as leaves.
x <- map_if(x, is_bare_list, recurse)


# Bad

# objects like data frames are treated as leaves
x <- map_if(x, is_bare_list, recurse)

# Objects like data frames are treated as leaves.
x <- map_if(x, is_bare_list, recurse)
```
