# Functions

## Naming

As well as following the general advice for [object names], strive to use verbs for function names:

```{r eval = FALSE}
# Good
add_row()
permute()

# Bad
row_adder()
permutation()
```

## Multi-line function defintions

There are two options if the function name and definition can't fit on a single line. In both cases, each argument goes on its own; the difference is how deep you indent it and where you put `)` and `{`:

*   **Single-indent**: indent the argument name with a single indent (i.e. two spaces).
    The trailing `)` and leading `{` go on a new line.

    ```{r, eval = FALSE}
    # Good
    long_function_name <- function(
      a = "a long argument",
      b = "another argument",
      c = "another long argument"
    ) {
      # As usual code is indented by two spaces.
    }
    ```

*   **Function-indent**: indent the argument name to match the opening `(` of `function`.
    The trailing `)` and leading `{` go on the same line as the last argument.

    ```{r, eval = FALSE}
    # Bad
    long_function_name <- function(a = "a long argument",
                                   b = "another argument",
                                   c = "another long argument") {
      # As usual code is indented by two spaces.
    }
    ```

These styles are designed to clearly separate the function definition from its body.

```{r, eval = FALSE}
# Bad
long_function_name <- function(a = "a long argument",
  b = "another argument",
  c = "another long argument") {
  # Here it's hard to spot where the definition ends and the
  # code begins, and to see all three function arguments
}
```

If a function argument can't fit on a single line, this is a sign you should rework the argument to keep it [short and sweet](https://design.tidyverse.org/defaults-short-and-sweet.html).

## S7

In S7, the method definition can be long because the function name is replaced by a method call that specifies the generic and dispatch classes. In this case we recommend the single-indent style.

```{r, eval = FALSE}
method(from_provider, list(openai_provider, class_any)) <- function(
  provider,
  x,
  ...,
  error_call = caller_env()
) {
  ...
}
```

If the method definition is too long to fit on one line, use the usual rules to
spread the method arguments across multiple lines:

```{r, eval = FALSE}
method(
  from_provider,
  list(openai_provider, class_any, a_very_long_class_name)
) <- function(
  provider,
  x,
  ...,
  error_call = caller_env()
) {
  ...
}
```

## `return()`

Only use `return()` for early returns. Otherwise, rely on R to return the result
of the last evaluated expression.

```{r eval = FALSE}
# Good
find_abs <- function(x) {
  if (x > 0) {
    return(x)
  }
  x * -1
}
add_two <- function(x, y) {
  x + y
}

# Bad
add_two <- function(x, y) {
  return(x + y)
}
```

Return statements should always be on their own line because they have important effects on the control flow. See also [inline statements](#inline-statements).

```{r, eval = FALSE}
# Good
find_abs <- function(x) {
  if (x > 0) {
    return(x)
  }
  x * -1
}

# Bad
find_abs <- function(x) {
  if (x > 0) return(x)
  x * -1
}
```

If your function is called primarily for its side-effects (like printing,
plotting, or saving to disk), it should return the first argument invisibly.
This makes it possible to use the function as part of a pipe. `print` methods
should usually do this, like this example from [httr](http://httr.r-lib.org/):

```{r eval = FALSE}
print.url <- function(x, ...) {
  cat("Url: ", build_url(x), "\n", sep = "")
  invisible(x)
}
```

## Comments

In code, use comments to explain the "why" not the "what" or "how". Each line
of a comment should begin with the comment symbol and a single space: `# `.

```{r, eval = FALSE}
# Good

# Objects like data frames are treated as leaves
x <- map_if(x, is_bare_list, recurse)


# Bad

# Recurse only with bare lists
x <- map_if(x, is_bare_list, recurse)
```

Comments should be in sentence case, and only end with a full stop if they
contain at least two sentences:

```{r, eval = FALSE}
# Good

# Objects like data frames are treated as leaves
x <- map_if(x, is_bare_list, recurse)

# Do not use `is.list()`. Objects like data frames must be treated
# as leaves.
x <- map_if(x, is_bare_list, recurse)


# Bad

# objects like data frames are treated as leaves
x <- map_if(x, is_bare_list, recurse)

# Objects like data frames are treated as leaves.
x <- map_if(x, is_bare_list, recurse)
```
