# ggplot2


## Introduction

Styling suggestions for `+` used to separate ggplot2 layers are very similar to those for `%>%` in pipelines.


## Whitespace

`+` should always have a space before it, and should be followed by a new line. This is true even if your plot has only two layers. After the first step, each line should be indented by two spaces.

If you are creating a ggplot off of a dplyr pipeline, there should only be one level of indentation.

```{r eval=FALSE}
# Good
iris %>%
  filter(Species == "setosa") %>%
  ggplot(aes(x = Sepal.Width, y = Sepal.Length)) +
  geom_point()

# Bad
iris %>%
  filter(Species == "setosa") %>%
  ggplot(aes(x = Sepal.Width, y = Sepal.Length)) +
    geom_point()

# Bad
iris %>%
  filter(Species == "setosa") %>%
  ggplot(aes(x = Sepal.Width, y = Sepal.Length)) + geom_point()
```


## Long lines

If the arguments to a ggplot2 layer don't all fit on one line, put each argument on its own line and indent:

```{r, eval = FALSE}
# Good
ggplot(aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() +
  labs(
    x = "Sepal width, in cm",
    y = "Sepal length, in cm",
    title = "Sepal length vs. width of irises"
  ) 

# Bad
ggplot(aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() +
  labs(x = "Sepal width, in cm", y = "Sepal length, in cm", title = "Sepal length vs. width of irises") 
```

ggplot2 allows you to do data manipulation, such as filtering or slicing, within the `data` argument. Avoid this, and instead do the data manipulation in a pipeline before starting plotting.

```{r eval=FALSE}
# Good
iris %>%
  filter(Species == "setosa") %>%
  ggplot(aes(x = Sepal.Width, y = Sepal.Length)) +
  geom_point()

# Bad
ggplot(filter(iris, Species == "setosa"), aes(x = Sepal.Width, y = Sepal.Length)) +
  geom_point()
```
