# Files {#package-files}

The majority of advice in @sec-files also applies to files in packages. Important differences are described below.

## Names

* If a file contains a single function, give the file the same name as the
  function.

* If a file contains multiple related functions, give it a concise,
  but evocative name.

* Deprecated functions should live in a file with `deprec-` prefix.

## Organisation

In a file that contains multiple functions, public functions and their
documentation should appear first, with private functions appearing after all
documented functions. If multiple public functions share the same documentation,
they should all immediately follow the documentation block.

See @sec-documentation for more thorough guidance on documenting functions
in packages.

```{r}
# Bad
help_compute <- function() {
  # ... Lots of code ...
}

#' My public function
#'
#' This is where the documentation of my function begins.
#' ...
#' @export
do_something_cool <- function() {
  # ... even more code ...
  help_compute()
}
```

```{r}
# Good
#' Lots of functions for doing something cool
#'
#' ... Complete documentation ...
#' @name something-cool
NULL

#' @describeIn something-cool Get the mean
#' @export
get_cool_mean <- function(x) {
  # ...
}

#' @describeIn something-cool Get the sum
#' @export
get_cool_sum <- function(x) {
  # ...
}
```
