# Pipes

## Introduction

Use `%>%` to emphasise a sequence of actions, rather than the object that the actions are being performed on. 

Avoid using the pipe when:

* You need to manipulate more than one object at a time. Reserve pipes for a 
  sequence of steps applied to one primary object.
 
* There are meaningful intermediate objects that could be given
  informative names.

## Whitespace

`%>%` should always have a space before it, and should usually be followed by a new line. After the first step, each line should be indented by two spaces. This structure makes it easier to add new steps (or rearrange existing steps) and harder to overlook a step.

```{r, eval = FALSE}
# Good
iris %>%
  group_by(Species) %>%
  summarize_if(is.numeric, mean) %>%
  ungroup() %>%
  gather(measure, value, -Species) %>%
  arrange(value)

# Bad
iris %>% group_by(Species) %>% summarize_all(mean) %>%
ungroup %>% gather(measure, value, -Species) %>%
arrange(value)
```

## Long lines

If the arguments to a function don't all fit on one line, put each argument on 
its own line and indent:

```{r, eval = FALSE}
iris %>%
  group_by(Species) %>%
  summarise(
    Sepal.Length = mean(Sepal.Length),
    Sepal.Width = mean(Sepal.Width),
    Species = n_distinct(Species)
  )
```

## Short pipes

It is ok to keep a one-step pipe in one line:

```{r, eval = FALSE}
# Good
iris %>% arrange(Species)

iris %>% 
  arrange(Species)

arrange(iris, Species)
```

Sometimes it's useful to include a short pipe as an argument to a function in a 
longer pipe. Carefully consider whether the code is more readable with a short 
inline pipe (which doesn't require a lookup elsewhere) or if it's better to move 
the code outside the pipe and give it an evocative name.

```{r, eval = FALSE}
# Good
x %>%
  select(a, b, w) %>%
  left_join(y %>% select(a, b, v), by = c("a", "b"))

x_join <-
  x %>%
  select(a, b, w)
y_join <-
  y %>%
  filter(!u) %>%
  gather(a, v, -b) %>%
  select(a, b, v)
left_join(x_join, y_join, by = c("a", "b"))

# Bad
x %>%
  select(a, b, w) %>%
  left_join(
    y %>% filter(!u) %>% gather(a, v, -b) %>% select(a, b, v), 
    by = c("a", "b")
  )
```

## No arguments

magrittr allows you to omit `()` on functions that don't have arguments. Avoid 
this.

```{r, eval = FALSE}
# Good
x %>% 
  unique() %>%
  sort()

# Bad
x %>% 
  unique %>%
  sort
```

## Assignment

Use a separate line for the target of the assignment followed by `<-`.

Personally, I think you should avoid using `->` to create an object at the end 
of the pipe. While starting with the assignment is a little more work when 
writing the code, it makes reading the code easier. This is because the name 
acts as a heading, which reminds you of the purpose of the pipe.

```{r, eval = FALSE}
# Good
iris_long <-
  iris %>%
  gather(measure, value, -Species) %>%
  arrange(-value)

# Bad
iris_long <- iris %>%
  gather(measure, value, -Species) %>%
  arrange(-value)

iris %>%
  gather(measure, value, -Species) %>%
  arrange(-value) ->
  iris_long
```
